Scalable Clustering of Genotype Information using MapReduce

### 2015-03-06 doc/publications/journal2014/paper.tex
updated sections

\title[MapReduce Clustering]{Scalable Clustering of Genotype Information using MapReduce}
\author[O'Brien \textit{et~al}]{Aidan R O'Brien\,$^{1}$, Fabian A Buske\,$^{2}$ and Denis C. Bauer\,$^1$\footnote{to whom correspondence should be addressed}}
\address{$^{1}$CSIRO, Digital Productivity Flagship, 11 Julius Av, 2113, Sydney, Australia\\
$^{2}$Cancer Epigenetics Program, Cancer Research Division, Kinghorn Cancer Centre, Garvan Institute of Medical Research, 384 Victoria St, 2010, Sydney, Australia}

@@

Increasing the input to include the first three chromosomes (\ThreePhasethree{}) improves the accuracy with an \ARI{} of 0.825 (Table~\ref{Tab:02}).
We are reaching a saturation point for k-means clustering here, as including all 22 autosomes (\FullPhasethree{}) improves the \ARI{} only marginally to 0.84 (Table~\ref{Tab:02}).
%DB is that really GB or should that be 6.1GB?
However, this demonstrates that performing clustering using whole-genome variants on research infrastructure is feasible using Mahout as the resource consumption lies within realistic boundaries. The time consumption increases to 25 hours and memory to 6144 GB (Table~\ref{Tab:02}).

@@

\subsection*{Result visualisation}
% DB: in the conclusion we discussed that we provide a "toolset to interpret Mahout's clustering output". We hence need to show the R-image for that, i.e. spit out distance matrix and read in with R to plot, e.g. doing Multidimensional Scaling to squash the matrix in 2D http://www.statmethods.net/advstats/mds.html

>>>>

\title[MapReduce Clustering]{Scalable Clustering of Genotype Information using MapReduce}
\author[O'Brien \textit{et~al}]{Aidan R O'Brien\,$^{1}$, Neil Saunders\,$^1$,Fabian A Buske\,$^{2}$ and Denis C. Bauer\,$^1$\footnote{to whom correspondence should be addressed}}
\address{$^{1}$CSIRO, Digital Productivity Flagship, 11 Julius Av, 2113, Sydney, Australia\\
$^{2}$Cancer Epigenetics Program, Cancer Research Division, Kinghorn Cancer Centre, Garvan Institute of Medical Research, 384 Victoria St, 2010, Sydney, Australia}

@@

Increasing the input to include the first three chromosomes (\ThreePhasethree{}) improves the accuracy with an \ARI{} of 0.825 (Table~\ref{Tab:02}).
We are reaching a saturation point for k-means clustering here, as including all 22 autosomes (\FullPhasethree{}) improves the \ARI{} only marginally to 0.84 (Table~\ref{Tab:02}).
%DB is that really GB or should that be 6.1GB?
However, this demonstrates that performing clustering using whole-genome variants on research infrastructure is feasible using Mahout as the resource consumption lies within realistic boundaries. The time consumption increases to 25 hours and memory to 6 GB (Table~\ref{Tab:02}).

@@

\subsection*{Result visualisation}
% DB: in the conclusion we discussed that we provide a "toolset to interpret Mahout's clustering output". 
% DB: can we discuss here how we get a visual representation from the clustering (ClusterDumper+Gephi) and what the image tells us..
